---
title: "Chapter 10 Body Fat Examples"
author: "Tian Zheng"
date: "December 1, 2015"
output: html_document
---
Use "bodyfat" dataset as an example to illustrate how to compute diagnostic measurements in R.

### Description of the "bodyfat" dataset:
- 1st column: Triceps Skinfold Thickness (X1)
- 2nd column: Thigh Circumference (X2)
- 3rd column: Midarm Circumference (X3)
- 4th Column: Body fat Y (X4)

```{r}
load("bodyfat.RData")
colnames(bodyfat) <- c("x1","x2","x3","y")
```

### 10.1 Added-Variable Plots
Added-variable plots consider the marginal role of a predictor variable $X_k$ given that the other predictors are already in the model. It is related to $SSR(X_k|X_1, ..., X_{k-1}, X_{k+1}, ...,
    X_{P-1})$ but contains more information.

```{r}
attach(bodyfat)
lm1 <- lm(y~x1+x2)
e1 <- lm(y~x1)$residual
e2 <- lm(y~x2)$residual
e12 <- lm(x1~x2)$residual
e21 <- lm(x2~x1)$residual
```

### Figure 10.4: Residual Plots and Added Variable Plots
```{r}
par(mfrow=c(2,2))
plot(x1,lm1$residual,xlab="X1",ylab="Residual",
     main="(a) Residual Plot agianst X1")
abline(h=0,lty=2)
plot(e12,e2,xlab="e(X1|X2)",ylab="e(Y|X2)",
     main="(b) Added-Variable Plot for X1")
abline(lm(e2~e12-1))
abline(h=0,lty=2)
plot(x2,lm1$residual,xlab="X2",ylab="Residual",
     main="(c) Residual Plot against X2")
abline(h=0,lty=2)
plot(e21,e1,xlab="e(X2|X1)",ylab="e(Y|X1)",
     main="(d) Added-Variable Plot for X2")
abline(lm(e1~e21-1))
abline(h=0,lty=2)
```

### 10.2 Studentized Deleted Residuals

Identifying outlying $Y$ observations using studentized deleted residuals: $t_i=\frac{d_i}{s(d_i)}=\frac{e_i}{\sqrt{MSE_{(i)}(1-h_{ii})}}$

Test for outliers: The studentized deleted
            outliers $t_i$'s will be compared with $t(1-\alpha/2n;
            n-p-1)$. It is a family of tests using Bonferroni
            procedure.
            
```{r}
library(MASS)
studres(lm1)  #studress gives the studentized deleted residuals
```

### 10.3 Hat Matrix Leverage Values
Identifying outlying $X$ observations using Hat Matrix Leverage values. 

Criterion for detecting outlying X observations: A
    leverage $h_{ii}$ is usually considered to be large if it is
    more than twice as large as the mean leverage value, denoted by
    $\bar{h}=\sum{h_{ii}}/n=p/n$, that $h_{ii} > 2p/n$.

#### Figure 10.7: Scatter Plot of Thigh Circiumference against Triceps Skinfold

```{r}
plot(x1,x2,type="n",xlab="Triceps Skinfold Thickness", 
     ylab="Thigh Circumference",
     main="Scatter Plot of X2 against X1 with Case Number Labeled")
text(x1,x2,1:20)
```

Here `hatvalues` gives the diagonal values of the hat matrix

```{r}
hatvalues(lm1)
```

### 10.4 DFFITS, Cook's Distance and DFBETAS Measures

Here, we consider identifying Influential Cases using  deleted-one-case regressions.

DFFITS measures the influence of a case on the fitted value of itself. For the $i^{th}$ observation, $(DFFITS)_i=\frac{\hat{Y}_i-\hat{Y}_{i(i)}}{\sqrt{MSE_{(i)}h_{ii}}}$.

Criterion:
    - $|(DFFITS)_i| > 1$ for small to medium data sets $n <
    30$;
    - $|(DFFITS)_i| > 2\sqrt{p/n}$ for large data sets.

```{r}
dffits(lm1)
```

Cook's distance measures the aggregate influence of a
    case on the $n$ fitted values. For the $i^{th}$ observation,
    \$D_i=\frac{\sum{(\hat{Y}_j-\hat{Y}_{j(i)})^2}}{p MSE}$.

    Criterion: If the corresponding percentile values from $F(p,
    n-p)$ for $D_i$,
    - approximately $10\% \sim 20\%$ or less, little
        influence.
    - $\sim 30-40\%$, some influence.
    - $ > 50\%$, major influence.
  
```{r}
mycook <- cooks.distance(lm1) 
mycook
```

DFBETAS measures the influence of a case on regression coefficients.

Let $c_{kk}$ be the $k^{th}$ diagonal term of $(X^TX)^{-1}$. For the $i^{th}$ observation and the $k^{th}$ coefficient $b_k$,
$(DFBETAS)_{k(i)}=\frac{b_k-b_{k(i)}}{\sqrt{MSE_{(i)}c_{kk}}}$.

Criterion:
- $|(DFBETAS)_{k(i)}| > 1$ for small to medium data sets $n <
    30$;
- $|(DFBETAS)_{k(i)}| > 2/\sqrt{n}$ for large data sets.

```{r}
dfbetas(lm1)
```

### Figure 10.8: Proportional Influence Plot and Index Influence Plot

```{r}
par(mfrow=c(1,2),pty="s")
plot(predict(lm1),lm1$residuals,
     cex=mycook/max(mycook)*4,pch=19,
     xlab="YHAT",ylab="Residual",
     main=" (a) Proportional Influcence Plot")
plot(1:20,mycook,pch=19,cex=0.5,
     xlab="Case Index Number",ylab="Cook's Distance D", 
     main=" (b) Index Influence Plot")
lines(1:20,mycook)
```

### 10.5 Variance Inflation Factor

```{r}
lm2 <- lm(y~x1+x2+x3)
library(car)
vif(lm2)
```

```{r}
detach(bodyfat)
```